---
title: "Apex"
author: "Bill Kayser"
date: "March 7, 2016"
output: ioslides_presentation
---

# Measuring User Experience with Apdex

## What is Apdex?

* An Apdex score is a standardized measure indicating whether your site is meeting it's obligations
* Values from 0 (total failure) to 100 (every request succeeds)
* Scored based on responses bucketed according to response time thresholds:
    * Satisfied
    * Tolerating
    * Failed
* Timeouts and errors automatically placed in Failed bin.

## What is Apdex?

The Apdex Score is determined by the formula: 

$$\frac{N_s + (1/2 * N_t)}{N}$$

Where:

* $N_s$ = The number of requests completed in no more than $T$ seconds
* $N_t$ = The number of requests completed between $T$ and $4T$
* $N_f$ = The number of requests taking longer than $4T$, including requests that had errors or timed out (ignored by the formula).
* $N$ = $N_s + N_t + N_f$ = the total number of requests processed.

$T$ is referred to as the Apdex T parameter and is set individually for each application based on what
is considered a satisfactory response time from a business perspective.

## Configuring Apdex T

```{r, cache=F}
ex3 <- read_transactions('./data/timeseries/news-site.rds', 'backend')
t_limits <- signif(quantile(ex3$events$value, c(0.2, 0.96)), 2)

apdex_example <- apdex_hist(ex3)

renderPlot({
  apdex_example(input$apdex_t)
})

sliderInput("apdex_t", label = "Apdex T:", width='100%', 
              min = t_limits[1], max = t_limits[2], value = signif(quantile(ex3$events$value, 0.5),2), step = 100)
```

## Apdex Buttons

Apdex Scores for a group of apps can be enumerated with guages, but a richer visual 
is the "Apdex Button".

```{r, cache=F, fig.height=2}
ex4 <- read_transactions('./data/timeseries/news-site.rds', 'backend')
t_limits <- signif(quantile(ex4$events$value, c(0.2, 0.96)), 2)

apdex_button(ex4, signif(seq(sqrt(t_limits[1]), sqrt(t_limits[2]), length.out = 6)^2, 2))
```

* Each application's status is represented with its own "Traffic Light"
* Green &rarr; Good
* Red &rarr; Bad

## Apdex Buttons


```{r}

buttons <- apdex_grid(app_group)
svec <- sapply(app_group, function(data) { as.integer(100 * data$summary$score) })
scoredata <- data.frame(score=svec, 
                        index=1:length(app_group),
                        q=findInterval(svec, c(82, 91)))
scores <- ggplot(scoredata) +
    geom_text(aes(x=0, y=index, color=as.factor(q), label=score), size=10) +
    scale_color_manual(values=c('red', '#666666', 'darkgreen')) +
    ylim(0.5, length(app_group)+0.5)+
    blank_theme

sparkdata <- lapply(app_group, function(data) {
    filter(data$plotlines, measure=='median') %>%
        mutate(name=data$name)
}) %>% bind_rows

sparklines <- ggplot(sparkdata) +
    aes(x=timestamp, y=value) +
    geom_line() +
    facet_grid(name ~ ., scales='free_y') +
    xlim(min(sparkdata$timestamp), min(sparkdata$timestamp + minutes(10))) +
    blank_theme +
    theme(panel.background = element_rect(fill='#EEEEEE'))

multiplot(sparklines, buttons, scores, cols=3, widths=c(40,6, 4))
```

## Apdex vs Mean

```{r}

plotdata <- dcast(ex3$plotlines, timestamp ~ measure, drop=F) %>%
    mutate(apdex.score=score*max(mean))

ggplot(plotdata) +
    aes(x=timestamp) +
    geom_line(aes(y=mean, color='mean', linetype='mean')) + 
    geom_line(aes(y=median, color='median', linetype='median')) + 
    geom_line(aes(y=apdex.score, color='score', linetype='score')) +
    summary_colorscale + summary_linescale +
    my_plot_theme

```

## Cost/Benefit - Apdex {.cvb}

```{r, fig.width=10, fig.height=6}
cvb <- cvb_add(cvb, 'Apdex', Efficiency=7, Value=8)
cvb_plot(cvb)
```
