---
title: "Faceted Views"
author: "Bill Kayser"
date: "March 7, 2016"
output: 
  ioslides_presentation: 
    css: ./styles.css
    keep_md: no
    logo: images/logo_text.png
    smaller: yes
    widescreen: yes
---

## Faceted Views

```{r init_sc, echo=F, message=F, warning=F}
source('init.R')
```

Looking at collection of related things

* Cluster of servers
* Application server nodes
* Pages in an application
* Response times broken down by browser
    
## Sparklines - Mean

```{r}
ex <- read_transactions('./data/timeseries/by_shard.rds', 'backend', 'shard', 8)

plotlines <- dcast(ex$plotlines, name + timestamp ~ measure, drop=F)
plotlines[is.na(plotlines$mean),'mean'] <- 0

z=qnorm(0.75)
max_t = quantile(ex$events$value, 0.95)

```

```{r}
ggplot(plotlines) +
    aes(x=timestamp, y=mean) +
    geom_line(aes(color='mean', linetype='mean', linejoin='round')) +
    blank_theme +
    theme(panel.background=element_rect(fill='#EEEEEE')) +
    summary_colorscale + summary_linescale +
    facet_grid(name ~ .) 
    
```

## Sparklines - Filled

```{r}
ggplot(plotlines) +
    aes(x=timestamp, y=mean) +
    geom_area(fill=region_fill, color='black') +
    blank_theme +
#    theme(panel.background=element_rect(fill='#EEEEEE')) +
    summary_colorscale + summary_linescale +
    facet_grid(name ~ .) 
    
```

## Sparklines - Inner Quartiles (est.)

```{r}
ggplot(plotlines) +
    aes(x=timestamp, ymax=gmean*(gsd^z),
        ymin=gmean/(gsd^z)) +
    geom_ribbon(fill=region_fill) +
    blank_theme +
    theme(panel.background=element_rect(fill='#EEEEEE')) +
    summary_colorscale + summary_linescale +
    facet_grid(name ~ .) #, scales='free_y') 
```


## Sparklines - Inner Quartiles, Free Y Axis

```{r}
ggplot(plotlines) +
    aes(x=timestamp, ymax=gmean*(gsd^z),
        ymin=gmean/(gsd^z)) +
    geom_ribbon(fill=region_fill) +
    blank_theme +
    theme(panel.background=element_rect(fill='#EEEEEE')) +
    summary_colorscale + summary_linescale +
    facet_grid(name ~ ., scales='free_y') 

```

## Sparklines - Horizon Plots

```{r}

hdata <- horizon_data(ex,n=4)

ggplot(hdata$df) +
    aes(x=timestamp, y=mean, fill=horizon) +
    scale_fill_brewer(palette='GnBu')+
    geom_area(position='identity', color='#999999') +
    geom_hline(aes(yintercept=hdata$horizon)) +
    geom_hline(aes(yintercept=0)) +    
    summary_colorscale + summary_linescale  +
    blank_theme +
#    theme(panel.background=element_rect(fill='#EEEEEE')) +
    facet_grid(name ~ ., scales='free_y') 

```

## Sparklines - Frequency Plots

```{r}
# ggplot(filter(events, value < max_t)) +
ggplot(filter(ex$events, value < max_t)) +
    aes(x=value) +
    geom_histogram(fill='#333333', bins=min(200, max_t)) +
    blank_theme +
    theme(panel.background=element_rect(fill='#EEEEEE')) +
    facet_grid(name ~ ., scales='free_y')

```

## Sparklines - Frequency + Rug

```{r}
ggplot(filter(ex$events, value < max_t)) +
    aes(x=value) +
    geom_histogram(fill='#333333', bins=min(200, max_t)) +
    blank_theme +
    theme(panel.background=element_rect(fill='#EEEEEE')) +
    facet_grid(name ~ ., scales='free_y') +
    geom_rug(sides='b')

```

## Sparklines - Density Filaments

```{r}
freq <- lapply(split(ex$events, ex$events$name), function(df) {
    d <- density(df$value, from=0, to=max_t)
    data.frame(t=d$x, d=d$y, name=df$name[1], stringsAsFactors = F)
}) %>% bind_rows %>% 
    tbl_df %>%
    mutate(name=as.factor(name))

ggplot(freq) +
    aes(x=t,y=1) +
    geom_raster(aes(fill=sqrt(d))) +
    scale_fill_continuous(low='#DDFFDD', high='#FF1111') +
    blank_theme +
    theme(panel.background=element_rect(fill='#EEEEEE')) +
    geom_vline(aes(xintercept=median), size=2, color='#666666', data=ex$summary) +
    facet_grid(name ~ ., scales='free_y')
```
