---
title: "APM Data and Visualization"
runtime: shiny
author: "Bill Kayser"
date: "February 16, 2016"
output: 
  ioslides_presentation: 
    css: ./styles.css
    keep_md: yes
    logo: images/logo_text.png
    smaller: yes
    widescreen: yes
---
```{r init, echo=F, message=F, warning=F}
source('init.R')
```

# In Graphic Detail | Data Science in Application Performance Management

Bill Kayser

Distinguished Engineer, New Relic

March 14, 2016


```{r, child='1._Intro.Rmd'}
```

```{r, child='2._Simple_Stats.Rmd'}
```

```{r, child='3._Geometric.Rmd'}
```

```{r, child='4._Apdex.Rmd'}
```


## Review - Summary Values

* For response time charts use geometric mean for an inexpensive measure.
* Forget the standard deviation, use the geometric standard deviation.
* For counts, throughput, use mean.
* Use Apdex for a user-centered metric.

# Additional Visualizations

## Box Plots

```{r boxplots, cache=F}
ex5 <- read_transactions('./data/timeseries/news-site.rds', 'backend')
boxplot <- mixed_charts(ex5)

renderPlot({
  boxplot(input)
})

radioButtons('bp', NULL, c(Scatterplot='scatter',Boxplot='boxplot'))

```

```{r, child='5._Spark_Charts.Rmd'}
```


## Throughput vs Response Time

```{r t_vs_rt}

files <- list.files('./data/scalability', '*.rds', full.names = T)
scalability_chart <- function(file) {
    df <- readRDS(file)
    ggplot(df) +
        aes(x=count, y=average) +
        geom_point(aes(color=hour),size=3, alpha=0.8)  +
        #geom_smooth() +
        scale_color_manual(values=div_gradient_pal(low='#007700', mid='#00FFFF', high='#992222')(seq(0, 1, length.out = 24))) +
        my_plot_theme + 
        theme(legend.position='none', text=element_blank(),
              panel.border=element_rect(fill=NA))
}

charts <- lapply(files, scalability_chart)
multiplot(plotlist=charts, cols=3)
    
```

Plots of throughput on the X axis and latency on the Y axis.

## Animation

Show Ward's video

## Data Science and APM {.build .smaller}

> * Quality visualizations give us greater visibility into the data.
> * Applying principles froms data science improves the fidelity of data and lowers the cost of acquisition.
> * Use both if you want to know your application

# Thanks...

## Bill Kayser

* @bravoking
* `github.com/bkayser`
* This talk: `github.com/bkayser/apmds`
* New Relic: www.newrelic.com
* R Library for getting New Relic Data: `github.com/bkayser/NewRelicR`

## References

Some articles, books, blogs, etc.

