---
title: "In Graphic Detail"
author: "Bill Kayser"
date: "March, 2016"
output:
  ioslides_presentation:
    css: ./styles.css
    keep_md: yes
    logo: images/logo_text.png
    smaller: yes
    widescreen: yes
  beamer_presentation: default
---
```{r init, echo=F, message=F, warning=F}
source('init.R')
```

# In Graphic Detail | Data Science in Application Performance Management

Bill Kayser

Distinguished Engineer, New Relic

March, 2016


```{r, child='1._Intro.Rmd'}
```

```{r, child='2._Simple_Stats.Rmd'}
```

```{r, child='3._Geometric.Rmd'}
```

```{r, child='4._Apdex.Rmd'}
```

```{r, child='5._Facets.Rmd'}
```

## Box Plots

```{r boxplots, cache=F}
data <- read_transactions('./data/timeseries/storefront-deploy.rds', 'frontend')
boxplot <- mixed_charts(data)
boxplot(list(bp='boxplot'))

```

## Throughput vs Response Time

```{r t_vs_rt}

files <- list.files('./data/scalability', '*.rds', full.names = T)
scalability_chart <- function(file) {
    df <- readRDS(file)
    ggplot(df) +
        aes(x=count, y=average) +
        geom_point(aes(color=hour),size=3, alpha=0.8)  +
        #geom_smooth() +
        scale_color_manual(values=div_gradient_pal(low='#007700', mid='#00FFFF', high='#992222')(seq(0, 1, length.out = 24))) +
        my_plot_theme + 
        theme(legend.position='none', text=element_blank(),
              panel.border=element_rect(fill=NA))
}

charts <- lapply(files, scalability_chart)
multiplot(plotlist=charts, cols=3)
    
```

Plots of throughput on the X axis and latency on the Y axis.  Color is time of day.

## Animation

[Traffic Animation](http://marlowe.datanerd.us/datasets/traffic.html)

## Summary - Monitoring Latency {.smaller}

<div class="columns-2">

### When you can collect a lot of data

* Use Scatterplots for deep visibility
* Use Median, Inner Quartiles for plots and sparkcharts
* Use histograms to identify multi-modal behavior
* Collect multiple dimensions for faceted views when possible


```{r, fig.width=4, fig.height=3.5}
cvb[,'Notice'] <- F
cvb_plot(cvb, scale=0.5)

```

</div>

## Summary - Monitoring Latency {.build .smaller}

### When space is a premium

* Be careful using the mean
* Use Geometric Mean to approximate the median
* Use Geometric Standard Deviation to approximate inner quartile regions
* Use Apdex for monitoring the quality of the customer experience

### Consider alternate visualizations

* Throughput vs Response Time Scatterplots
* Density Filaments
* Horizon Plots
* Animation

## Thanks!

Bill Kayser (@bravoking) - www.newrelic.com

* This talk: [bkayser.github.io/apmviz](http://bkayser.github.io/apmviz)
* Source for this talk: [github.com/bkayser/apmviz](http://github.com/bkayser/apmviz)

* D3/Javascript version live demo: http://marlowe.datanerd.us
* D3/Javascript source: [github.com/newrelic/marlowe](http://github.com/newrelic/marlowe)

* R Library for getting New Relic Data: `github.com/bkayser/NewRelicR`


